---
title: "Winson code (YFV allison ver.)"
author: "Winson (allison edit)"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#Loading packages 
library(ggplot2)
library(dplyr)
library(tidyverse)
library(lubridate)
library(Polychrome)
library(sf)
library(here)
library(zoo)
library(tidytext)
library(scales)
library(geobr)
library(RColorBrewer)
library(foreign)

setwd("/Users/allison/Desktop/Research_BentoLab/SP26")
```

This script produces the following outputs:
1. Number of Infections
~~  - by Age and State ~~ (no age data)
~~  - by Age and Sex ~~
~~  - by Sex ~~
~~  - by Ethnicity and Sex~~
~~  - by Zone and Sex ~~
2. Annual Incidence
  - by Year and State
  - by Municipality and State
3. Time Series Incidence
  - Monthly Incidence per State
    - Log Scale
    - 3 month MA Log Scale
    - 6 month MA Log Scale
    - 3 month MA with 95% CI Log Scale
    - 6 month MA with 95% CI Log Scale
4. Incidence by Age
  - Age and Sex
  - Age grouped by State
5. Spatial Representations
  - Municipality Level
  - State Level


```{r}
#Loads the YFV Dataset
YFV <- read.csv("data/municipality_incidence_yellow_fever_William_dataset.csv")

#Loads the Dataset with Brazil's size per age, year, state and municipality
population_size_2010_2024 <- read.csv("data/population_size_2010_2024.csv")

```


```{r calculate incidence for state, echo=FALSE}

# 1: Group cases by state and year
YFV_state <- YFV %>%
  group_by(state, year) %>%
  summarise(cases = sum(total_cases, na.rm = T),
            population = sum(total_population, na.rm = T),
            incidence_100k = (cases / population) * 100000,
            .groups = "drop")

# Remove zero-incidence rows (for log scale)
YFV_state_cleaned <- YFV_state %>%
  filter(incidence_100k > 0)

# Step 3. Create a lookup table for IBGE codes -> state abbreviations
state_lookup <- c(
  "11" = "RO", "12" = "AC", "13" = "AM", "14" = "RR", "15" = "PA",
  "16" = "AP", "17" = "TO", "21" = "MA", "22" = "PI", "23" = "CE",
  "24" = "RN", "25" = "PB", "26" = "PE", "27" = "AL", "28" = "SE",
  "29" = "BA", "31" = "MG", "32" = "ES", "33" = "RJ", "35" = "SP",
  "41" = "PR", "42" = "SC", "43" = "RS", "50" = "MS", "51" = "MT",
  "52" = "GO", "53" = "DF"
)

# Step 4. Prepare population dataset
# - Ensure municipality codes are 6-digit strings
# - Extract first 2 digits as state code
# - Map to state abbreviation
population_size_2010_2024 <- population_size_2010_2024 %>%
  mutate(
    municipality_code_age = sprintf("%06d", as.integer(municipality_code_age)),
    state_code = substr(municipality_code_age, 1, 2),
    state = state_lookup[state_code]
  )

# Step 5. Aggregate population to state-year level
population_state <- population_size_2010_2024 %>%
  group_by(state, year) %>%
  summarise(Total_pop = sum(Total, na.rm = TRUE), .groups = "drop")

# Step 6. Join case counts with population and calculate incidence per 100,000
incidence_per_state <- cases_state %>%
  left_join(population_state, by = c("year","state")) %>%
  mutate(incidence_per_state_100k = (cases / Total_pop) * 100000)

# Step 7. Remove rows with missing incidence (NAs from unmatched state-year combinations)
incidence_per_state_clean <- incidence_per_state %>%
  filter(!is.na(incidence_per_state_100k))

# Step 8. Plot heatmap
heatmap_plot_1 <- ggplot(
  incidence_per_state_clean,
  aes(
    x = factor(year),
    y = reorder(state, incidence_per_state_100k, FUN = max),
    fill = incidence_per_state_100k
  )
) +
  geom_tile(color = "#FCF0CE") +
  scale_fill_gradient(
    low = "#FCF0CE",
    high = "#D4180A",
    trans = "log",  
    breaks = scales::trans_breaks("log10", function(x) 10^x),
    labels = scales::trans_format("log10", scales::math_format(10^.x))
  ) +
  labs(
    title = "",
    x = "Year",
    y = "State",
    fill = "Chikv\nincidence\n"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid = element_blank()
  )
print(heatmap_plot_1)

ggsave(
  here::here("figures", "incidence_per_state.png"),
  plot = heatmap_plot_1,
  width = 10,
  height = 6,
  dpi = 300
)
```

```{r calculate incidence for municipalities, echo=FALSE}
# Prepare dataset with year and month extracted from collection_date
chik4 <- chik3 %>%
  mutate(municipality_code = sprintf("%06d", as.integer(Municipality_Code)) # standardize code
  )
# Summarize the number of cases by year and municipality
cases_per_muni <- chik4 %>%
  group_by(year, municipality_code) %>%
  summarise(cases = n(), .groups = "drop")
# Select total population by year and municipality
pop_per_muni <- population_size_2010_2024 %>%
  select(year, municipality_code_age, Total) %>%
  rename(municipality_code = municipality_code_age,
         population = Total)
# Join the two datasets
incidence_data_per_municipalities <- cases_per_muni %>%
  inner_join(pop_per_muni, by = c("year", "municipality_code"))

# Calculate incidence per 100,000 inhabitants
incidence_per_municipalities <- incidence_data_per_municipalities %>%
  mutate(incidence_per_municipalities_100k = (cases / population) * 100000)

# Preview the first results
head(incidence_per_municipalities)
# Remove rows with missing incidence (NAs from unmatched municipality-year combinations)
incidence_per_municipalities_clean <- incidence_per_municipalities %>%
  filter(!is.na(incidence_per_municipalities_100k))

incidence_per_muni <- incidence_per_municipalities_clean %>%
  mutate(state_code = substr(municipality_code, 1, 2))

#Create lookup table for state abbreviations
state_lookup <- c(
  "11" = "RO", "12" = "AC", "13" = "AM", "14" = "RR", "15" = "PA",
  "16" = "AP", "17" = "TO", "21" = "MA", "22" = "PI", "23" = "CE",
  "24" = "RN", "25" = "PB", "26" = "PE", "27" = "AL", "28" = "SE",
  "29" = "BA", "31" = "MG", "32" = "ES", "33" = "RJ", "35" = "SP",
  "41" = "PR", "42" = "SC", "43" = "RS", "50" = "MS", "51" = "MT",
  "52" = "GO", "53" = "DF"
)

#Replace state codes with abbreviations
incidence_per_muni <- incidence_per_muni %>%
  mutate(state = state_lookup[state_code])

# Create an order of states based on their maximum incidence
state_order <- incidence_per_state_clean %>%
  group_by(state) %>%
  summarise(max_incidence = max(incidence_per_state_100k, na.rm = TRUE)) %>%
  arrange(desc(max_incidence)) %>%
  pull(state)

# Define the custom order for states
state_order <- c("RJ", "ES", "MG", "SP", "PA", "PR",
                 "AP", "RR", "SC", "PI", "RS", "AM", 
                 "MS", "DF", "GO", "TO", "MT", "AL", 
                 "RO", "BA", "PE")

# Apply this order to the state variable
incidence_per_muni <- incidence_per_muni %>%
  mutate(state = factor(state, levels = state_order))

# Plot heatmap with custom facet order
heatmap_plot_2<-ggplot(incidence_per_muni,
       aes(x = factor(year),
           y = reorder_within(municipality_code, 
                              incidence_per_municipalities_100k, 
                              state),           
           fill = incidence_per_municipalities_100k)) +
  geom_tile(color = "#fcf0ce") +
  scale_fill_gradient(low = "#fcf0ce", high = "#d4180a") +
  facet_wrap(~ state, scales = "free_y") +
  scale_fill_gradient(
    low = "#FCF0CE",
    high = "#D4180A",
    trans = "log",  
    breaks = scales::trans_breaks("log10", function(x) 10^x),
    labels = scales::trans_format("log10", scales::math_format(10^.x))
  ) +
  labs(
    title = " ",
    x = "Year",
    y = "Municipality", 
    fill = "ChikV incidence per municipality (2010–2024) (per 100k)"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    panel.grid = element_blank(),
    strip.text = element_text(face = "bold"),
    legend.position = "bottom"
  )

print(heatmap_plot_2)

ggsave(
  here::here("figures", "incidence_heatmap_municipality.png"),
  plot = heatmap_plot_2,
  width = 8,
  height = 10,
  dpi = 300
)
```

```{r}
chik6 <- chik %>%
  mutate(
    year = lubridate::year(Collection_Date),
    month = lubridate::month(Collection_Date)
  )

# --- Step 2: Count ChikV cases per state-year-month ---
cases_state <- chik6 %>%
  group_by(state, year, month) %>%
  summarise(cases = n(), .groups = "drop")

# --- Step 3: Aggregate population to state-year-month level ---
population_state <- population_size_2010_2024 %>%
  mutate(
    municipality_code_age = sprintf("%06d", as.integer(municipality_code_age)),
    state_code = substr(municipality_code_age, 1, 2),
    state = state_lookup[state_code]
  ) %>%
  group_by(state, year) %>%
  summarise(Total_pop = sum(Total, na.rm = TRUE), .groups = "drop") %>%
  tidyr::crossing(month = 1:12)

# --- Step 4: Join case counts with population and compute incidence ---
incidence_per_state_order <- cases_state %>%
  left_join(population_state, by = c("year", "state", "month")) %>%
  mutate(
    incidence_per_state_100k = (cases / Total_pop) * 100000 + 0.001,  # add small offset to avoid log(0)
    year_month = lubridate::ymd(paste(year, month, "01", sep = "-"))
  ) %>%
  arrange(state, year_month) %>%
  filter(!is.na(incidence_per_state_100k))

# --- Step 5: Base time series plot (log scale) ---
move_plot <- ggplot(incidence_per_state_order, aes(x = year_month, y = incidence_per_state_100k)) +
  geom_line(color = "grey60", alpha = 0.6) +
  geom_smooth(method = "loess", span = 0.2, se = TRUE, aes(color = state, fill = state)) +
  facet_wrap(~state, scales = "free_y") +
  scale_y_log10(
    labels = comma_format(accuracy = 1),
    breaks = log_breaks(n = 6)
  ) +
  labs(
    title = "Monthly ChikV Incidence per State (Log Scale)",
    x = "Year",
    y = "Log^10(Incidence per 100,000)"
  ) +
  theme_minimal() +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.text = element_text(size = 10, face = "bold")
  )
print(move_plot)

# --- Step 6: 3-month moving average ---
incidence_ma_3m <- incidence_per_state_order %>%
  group_by(state) %>%
  arrange(year_month) %>%
  mutate(
    incidence_ma_3m = zoo::rollmean(incidence_per_state_100k, k = 3, fill = NA, align = "right")
  ) %>%
  ungroup() %>%
  filter(!is.na(incidence_ma_3m))

ma_ci_plot_3m <- ggplot(incidence_ma_3m, aes(x = year_month, y = incidence_ma_3m)) +
  geom_line(color = "grey50", alpha = 0.7) +
  geom_smooth(method = "loess", span = 0.2, se = TRUE, aes(color = state, fill = state)) +
  facet_wrap(~state, scales = "free_y") +
  scale_y_log10(
    labels = comma_format(accuracy = 1),
    breaks = log_breaks(n = 6)
  ) +
  labs(
    title = "Monthly ChikV Incidence per State (3-Month Moving Average, Log Scale)",
    x = "Year",
    y = "Log^10(Incidence per 100,000, 3M MA)"
  ) +
  theme_minimal() +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.text = element_text(size = 10, face = "bold")
  )
print(ma_ci_plot_3m)

# --- Step 7: 6-month moving average ---
incidence_ma_6m <- incidence_per_state_order %>%
  group_by(state) %>%
  arrange(year_month) %>%
  mutate(
    incidence_ma_6m = zoo::rollmean(incidence_per_state_100k, k = 6, fill = NA, align = "right")
  ) %>%
  ungroup() %>%
  filter(!is.na(incidence_ma_6m))

ma_ci_plot_6m <- ggplot(incidence_ma_6m, aes(x = year_month, y = incidence_ma_6m)) +
  geom_line(color = "grey50", alpha = 0.7) +
  geom_smooth(method = "loess", span = 0.2, se = TRUE, aes(color = state, fill = state)) +
  facet_wrap(~state, scales = "free_y") +
  scale_y_log10(
    labels = comma_format(accuracy = 1),
    breaks = log_breaks(n = 6)
  ) +
  labs(
    title = "Monthly ChikV Incidence per State (6-Month Moving Average, Log Scale)",
    x = "Year",
    y = "Log₁₀(Incidence per 100,000, 6M MA)"
  ) +
  theme_minimal() +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.text = element_text(size = 10, face = "bold")
  )
print(ma_ci_plot_6m)
```

```{r}
chik7 <- chik %>%
  mutate(
    year = year(Collection_Date),
    month = month(Collection_Date),
    municipality_code = sprintf("%06d", as.integer(Municipality_Code))
  )

cases_per_muni <- chik7 %>%
  group_by(municipality_code, year, month) %>%
  summarise(cases = n(), .groups = "drop")

pop_per_muni <- population_size_2010_2024 %>%
  select(year, municipality_code_age, Total) %>%
  rename(municipality_code = municipality_code_age,
         population = Total) %>%
  crossing(month = 1:12)

incidence_per_muni_monthly <- cases_per_muni %>%
  inner_join(pop_per_muni, by = c("year", "municipality_code", "month")) %>%
  mutate(
    incidence_per_muni_100k = (cases / population) * 100000,
    incidence_per_muni_100k = incidence_per_muni_100k + 0.001,  # avoid log(0)
    year_month = ymd(paste(year, month, "01", sep = "-")),
    state_code = substr(municipality_code, 1, 2),
    state = state_lookup[state_code]
  ) %>%
  filter(!is.na(incidence_per_muni_100k))
plot_ma_municipality <- function(df, k, ma_label) {
  
  ma_df <- df %>%
    arrange(municipality_code, year_month) %>%
    group_by(municipality_code) %>%
    mutate(ma = zoo::rollmean(incidence_per_muni_100k, k = k, fill = NA, align = "right")) %>%
    ungroup() %>%
    filter(!is.na(ma)) %>%
    group_by(municipality_code) %>%
    filter(n() >= 12) %>% 
    ungroup()
  
  ggplot(ma_df, aes(x = year_month, y = ma, group = municipality_code)) +
    geom_line(color = "grey60", alpha = 0.5) +
    geom_smooth(method = "loess", span = 0.5, se = TRUE, aes(color = state, fill = state)) +
    facet_wrap(~state, scales = "free_y") +
    scale_y_log10(                                     # ← log scale here
      labels = scales::comma_format(accuracy = 1),
      breaks = scales::log_breaks(n = 5)
    ) +
    labs(
      title = paste("Monthly ChikV Incidence per Municipality (", ma_label, "-Month MA with 95% CI, Log Scale)", sep = ""),
      x = "Year",
      y = paste("Log₁₀(Incidence per 100,000, ", ma_label, "M MA)", sep = "")
    ) +
    theme_minimal() +
    theme(
      legend.position = "none",
      axis.text.x = element_text(angle = 45, hjust = 1),
      strip.text = element_text(size = 10, face = "bold")
    )
}

# Plot 3-month MA (log scale)
ma_ci_plot_3m_muni <- plot_ma_municipality(incidence_per_muni_monthly, k = 3, ma_label = "3")
print(ma_ci_plot_3m_muni)

# Plot 6-month MA (log scale)
ma_ci_plot_6m_muni <- plot_ma_municipality(incidence_per_muni_monthly, k = 6, ma_label = "6")
print(ma_ci_plot_6m_muni)
```

```{r}
chik_sex_age <- chik %>%
  mutate(
    age_group = cut(
      age_years,
      breaks = seq(0, 100, by = 10),
      labels = paste(seq(0, 95, by = 10), seq(4, 99, by = 10), sep = "-"),
      include.lowest = TRUE, right = TRUE
    )
  ) %>%
  filter(!is.na(Sex_Code), !is.na(age_group)) %>%
  group_by(age_group, Sex_Code) %>%
  summarise(cases = n(), .groups = "drop")


total_pop <- population_size_2010_2024 %>%
  summarise(population = sum(Total, na.rm = TRUE)) %>%
  pull(population)

incidence_sex_age <- chik_sex_age %>%
  mutate(
    population = total_pop,  # same for all groups
    incidence_per_100k = (cases / population) * 100000,
    se = sqrt(cases) / population * 100000,
    ci_lower = pmax(0, incidence_per_100k - 1.96 * se),
    ci_upper = incidence_per_100k + 1.96 * se
  )

# Step 4: Plot
ggplot(incidence_sex_age, aes(x = age_group, y = incidence_per_100k, fill = Sex_Code)) +
  geom_col(position = position_dodge(width = 0.9), width = 0.8) +
  geom_errorbar(
    aes(ymin = ci_lower, ymax = ci_upper),
    position = position_dodge(width = 0.9),
    width = 0.3
  ) +
  scale_fill_manual(
    values = c("F" = "#e41a1c", "M" = "#377eb8"),
    name = "Sex",
    labels = c("Female", "Male")
  ) +
  labs(
    title = "Incidence for Chikungunya Cases Grouped by Age per 100,000",
    x = "Age Group (Years)",
    y = "Approximate Incidence per 100,000"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    legend.position = "top",
    axis.text.x = element_text(angle = 45, hjust = 1)
  )


```

```{r}
chik_sex_age_state <- chik %>%
  mutate(
    state_code = substr(Municipality_Code, 1, 2),  # extract state code
    age_group = cut(
      age_years,
      breaks = seq(0, 100, by = 10),
      labels = paste(seq(0, 95, by = 10), seq(4, 99, by = 10), sep = "-"),
      include.lowest = TRUE, right = TRUE
    )
  ) %>%
  filter(!is.na(Sex_Code), !is.na(age_group), !is.na(state_code)) %>%
  group_by(state_code, age_group, Sex_Code) %>%
  summarise(cases = n(), .groups = "drop")

# Step 2: Aggregate total population by state (no age breakdown)
population_state <- population_size_2010_2024 %>%
  mutate(state_code = substr(municipality_code_age, 1, 2)) %>%
  group_by(state_code) %>%
  summarise(population = sum(Total, na.rm = TRUE), .groups = "drop")

# Step 3: Join and compute approximate incidence and 95% CI
incidence_sex_age_state <- chik_sex_age_state %>%
  left_join(population_state, by = "state_code") %>%
  mutate(
    incidence_per_100k = (cases / population) * 100000,
    se = sqrt(cases) / population * 100000,
    ci_lower = pmax(0, incidence_per_100k - 1.96 * se),
    ci_upper = incidence_per_100k + 1.96 * se
  )

# Step 4: Plot (faceted by state)
ggplot(incidence_sex_age_state, aes(x = age_group, y = incidence_per_100k, fill = Sex_Code)) +
  geom_col(position = position_dodge(width = 0.9), width = 0.8) +
  geom_errorbar(
    aes(ymin = ci_lower, ymax = ci_upper),
    position = position_dodge(width = 0.9),
    width = 0.3
  ) +
  scale_fill_manual(
    values = c("F" = "#e41a1c", "M" = "#377eb8"),
    name = "Sex",
    labels = c("Female", "Male")
  ) +
  facet_wrap(~ state_code, scales = "free_y") +
  labs(
    title = " Chikungunya Incidence per 100,000 for Age grouped by State",
    x = "Age Group (Years)",
    y = "Approximate Incidence per 100,000"
  ) +
  theme_minimal(base_size = 13) +
theme(
  legend.position = "top",
  axis.text.x = element_text(angle = 45, hjust = 1, size = 8),  # smaller x-axis labels
  axis.text.y = element_text(size = 9),                         # smaller y-axis labels
  strip.text = element_text(size = 9),                          # smaller facet labels
  axis.title = element_text(size = 11),                         # optional: slightly smaller axis titles
  panel.grid.major.x = element_blank()
) +
  theme(
    legend.position = "top",
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.major.x = element_blank()
  )
```

```{r}
library(data.table)
library(ggplot2)
library(geofacet)
library(dplyr)
library(forcats)
library(cowplot)
library(sf)
library(ISOweek)
library(ggpubr)
library(ggrepel)
library(beepr)
library(sjPlot)
library(stringr)
library(ggspatial)

## source functions
source("Brazil-arbovirus-meteo-main/R/plotting_functions.R") 
source("Brazil-arbovirus-meteo-main/R/extra_functions.R") 
source("Brazil-arbovirus-meteo-main/R/univariable_analysis_functions.R") 
source("Brazil-arbovirus-meteo-main/R/multivariable_analysis_functions.R") 
## read in shapefile 
shapefile <- readRDS("Brazil-arbovirus-meteo-main/data/Brazil_2_geobr.RDS") 


## read in shapefile
shapefile <- shapefile %>%
  mutate(code_muni = as.character(code_muni))
shapefile <- shapefile %>%
  mutate(code_muni = substr(code_muni, 1, nchar(code_muni) - 1))
incidence_per_muni <- incidence_per_muni %>%
  mutate(municipality_code = as.character(municipality_code))

#Join shapefile with incidence data by municipality code
shapefile_incidence <- shapefile %>%
  left_join(
    incidence_per_muni %>%
      select(municipality_code, incidence_per_municipalities_100k),
    by = c("code_muni" = "municipality_code")
  )

## path to read in graph for INLA models
GRAPH <- "data/Brazil2.graph"

library(sf)
library(ggplot2)

# Assuming your shapefile is already read as `shapefile`
# Example: shapefile <- st_read("path/to/your/shapefile.shp")

# Check the structure
glimpse(shapefile_incidence)


shapefile_projected <- st_transform(shapefile_incidence, 5880)

# 2. Plot with accurate scale bar
p <- ggplot(data = shapefile_projected) +
  # 1️⃣ Plot municipalities with fill based on incidence
  geom_sf(aes(fill = incidence_per_municipalities_100k),
          color = "grey40", size = 0.1) +
  
  # 2️⃣ Color scale with log10 transformation
  scale_fill_viridis_c(
    option = "plasma",
    na.value = "lightgrey",
    trans = "log10",
    name = "Incidence per 100k (log10)"
  ) +
  
  # 3️⃣ Coordinate system
  coord_sf(datum = NA, expand = FALSE) +
  
  # 4️⃣ Minimal theme and legend formatting
  theme_void() +
  theme(
    legend.position = "right",
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 9)
  ) +
  
  # 5️⃣ Add scale bar
  annotation_scale(
    location = "bl",          # bottom-left corner
    width_hint = 0.25,
    style = "ticks",
    text_cex = 0.8,
    line_width = 0.8,
    unit_category = "metric",
    bar_cols = c("black", "white"),
    text_pad = unit(0.1, "cm")
  ) +
  
  # 6️⃣ Add north arrow
  annotation_north_arrow(
    location = "bl",
    which_north = "true",
    pad_x = unit(0.3, "in"),
    pad_y = unit(0.8, "in"),
    style = north_arrow_fancy_orienteering
  )
p



```
```{r}
library(data.table)
library(ggplot2)
library(geofacet)
library(dplyr)
library(forcats)
library(cowplot)
library(sf)
library(ISOweek)
library(ggpubr)
library(ggrepel)
library(beepr)
library(sjPlot)
library(stringr)
library(ggspatial)

## source functions
source("Brazil-arbovirus-meteo-main/R/plotting_functions.R")
source("Brazil-arbovirus-meteo-main/R/extra_functions.R")
source("Brazil-arbovirus-meteo-main/R/univariable_analysis_functions.R")
source("Brazil-arbovirus-meteo-main/R/multivariable_analysis_functions.R")
## read in shapefile
shapefile_state <- readRDS("Brazil-arbovirus-meteo-main/data/Brazil_2_geobr.RDS")

incidence_per_state <- incidence_per_state %>%
  group_by(state) %>%
  mutate(incidence_avr = mean(incidence_per_state_100k, na.rm = TRUE)) %>%
  ungroup()

shapefile_state <- shapefile_state %>%
  mutate(abbrev_state = as.character(abbrev_state))

incidence_per_state <- incidence_per_state %>%
  mutate(state = as.character(state))

# --- Join shapefile and incidence data ---
shapefile_incidence_state <- shapefile_state %>%
  left_join(
    incidence_per_state %>% select(state, incidence_avr),
    by = c("abbrev_state" = "state")
  )

# --- Transform CRS for accuracy ---
shapefile_projected <- st_transform(shapefile_incidence_state, 5880)

# --- FIXED: Use the automatic geometry combination in summarise ---
shapefile_state_combined <- shapefile_projected %>%
  group_by(abbrev_state) %>%
  summarise(incidence_avr = first(incidence_avr)) %>%  # sf automatically combines geometry
  ungroup() %>%
  st_make_valid()

# --- FIXED: Simple centroid calculation ---
state_centroids <- shapefile_state_combined %>%
  st_centroid() %>%
  mutate(
    X = st_coordinates(.)[, 1],
    Y = st_coordinates(.)[, 2]
  ) %>%
  st_drop_geometry()

# --- FIXED: Clean map plot with NO white spots ---
p <- ggplot() +
  # Single layer with colored states - NO base layer
  geom_sf(
    data = shapefile_state_combined,
    aes(fill = incidence_avr),
    color = "black",  # Black borders to define states
    linewidth = 0.2   # Thin borders
  ) +
  # Add state labels
  geom_text(
    data = state_centroids,
    aes(x = X, y = Y, label = abbrev_state),
    size = 2.5,
    color = "white",
    fontface = "bold"
  ) +
  # Color scale
  scale_fill_viridis_c(
    option = "plasma",
    na.value = "gray80",
    trans = "log10",
    name = "Incidence per 100k (log10)",
    breaks = c(1, 3, 10, 30),
    labels = c("1.0", "3.0", "10.0", "30.0")
  ) +
  labs(title = "Average Chikungunya Incidence per 100k by State") +
  annotation_scale(
    location = "bl",
    width_hint = 0.25,
    style = "ticks",
    text_cex = 0.8,
    line_width = 0.8,
    unit_category = "metric",
    bar_cols = c("black", "white"),
    text_pad = unit(0.1, "cm")
  ) +
  annotation_north_arrow(
    location = "bl",
    which_north = "true",
    pad_x = unit(0.3, "in"),
    pad_y = unit(0.8, "in"),
    style = north_arrow_fancy_orienteering
  ) +
  theme_void() +
  theme(
    legend.position = "right",
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 8),
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    plot.background = element_rect(fill = "white", color = NA)
  )

p



```


```{r}
library(data.table)
library(ggplot2)
library(geofacet)
library(dplyr)
library(forcats)
library(cowplot)
library(sf)
library(ISOweek)
library(ggpubr)
library(ggrepel)
library(beepr)
library(sjPlot)
library(stringr)
library(ggspatial)

## source functions
source("Brazil-arbovirus-meteo-main/R/plotting_functions.R")
source("Brazil-arbovirus-meteo-main/R/extra_functions.R")
source("Brazil-arbovirus-meteo-main/R/univariable_analysis_functions.R")
source("Brazil-arbovirus-meteo-main/R/multivariable_analysis_functions.R")
## read in shapefile
shapefile_state <- readRDS("Brazil-arbovirus-meteo-main/data/Brazil_2_geobr.RDS")

# First, check the actual structure of the shapefile
print("Column names in shapefile:")
print(names(shapefile_state))

print("Geometry column name:")
print(attr(shapefile_state, "sf_column"))

incidence_per_state <- incidence_per_state %>%
  group_by(state) %>%
  mutate(incidence_avr = mean(incidence_per_state_100k, na.rm = TRUE)) %>%
  ungroup()

shapefile_state <- shapefile_state %>%
  mutate(abbrev_state = as.character(abbrev_state))

incidence_per_state <- incidence_per_state %>%
  mutate(state = as.character(state))

# --- Join shapefile and incidence data ---
shapefile_incidence_state <- shapefile_state %>%
  left_join(
    incidence_per_state %>% select(state, incidence_avr),
    by = c("abbrev_state" = "state")
  )

# --- Transform CRS for accuracy ---
shapefile_projected <- st_transform(shapefile_incidence_state, 5880)

# --- FIXED: Let sf automatically handle geometry combination ---
shapefile_state_combined <- shapefile_projected %>%
  group_by(abbrev_state) %>%
  summarise(incidence_avr = first(incidence_avr)) %>%  # sf automatically combines geometry
  ungroup() %>%
  st_make_valid()

# Check if we still have multiple polygons per state
print(paste("Number of polygons after summarise:", nrow(shapefile_state_combined)))

# If we still have multiple polygons, use st_union with the correct geometry column name
if (nrow(shapefile_state_combined) > 27) {  # Brazil has 27 states
  geometry_col <- attr(shapefile_state_combined, "sf_column")
  shapefile_state_combined <- shapefile_state_combined %>%
    group_by(abbrev_state) %>%
    summarise(
      incidence_avr = first(incidence_avr),
      geometry = st_union(!!sym(geometry_col))  # Use the actual geometry column name
    ) %>%
    ungroup() %>%
    st_make_valid()
}

# --- Create centroids for labeling ---
state_centroids <- shapefile_state_combined %>%
  st_centroid() %>%
  st_as_sf()

# --- Map with ONE label per state ---
p <- ggplot() +
  # Single layer with colored states
  geom_sf(
    data = shapefile_state_combined,
    aes(fill = incidence_avr),
    color = NA,  # No borders
    linewidth = 0
  ) +
  # Add ONE label per state using centroids
  geom_sf_text(
    data = state_centroids,
    aes(label = abbrev_state),
    size = 3,
    color = "white",
    fontface = "bold"
  ) +
  # Color scale
  scale_fill_viridis_c(
    option = "plasma",
    na.value = "gray80",
    trans = "log10",
    name = "Incidence per 100k (log10)",
    breaks = c(1, 3, 10, 30),
    labels = c("1.0", "3.0", "10.0", "30.0")
  ) +
  labs(title = "Average Chikungunya Incidence per 100k by State") +
  annotation_scale(
    location = "bl",
    width_hint = 0.25,
    style = "ticks",
    text_cex = 0.8,
    line_width = 0.8,
    unit_category = "metric",
    bar_cols = c("black", "white"),
    text_pad = unit(0.1, "cm")
  ) +
  annotation_north_arrow(
    location = "bl",
    which_north = "true",
    pad_x = unit(0.3, "in"),
    pad_y = unit(0.8, "in"),
    style = north_arrow_fancy_orienteering
  ) +
  theme_void() +
  theme(
    legend.position = "right",
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 8),
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    plot.background = element_rect(fill = "white", color = NA)
  )

p
```



